Probability of causal contact between interstellar civilizations through MonteCarlo simulations

### 2019-12-10 manuscript.tex
descripcion de la figura 1

Some complex stochastic processes can be efficiently modeled with the
discrete--event (DE) simulation approach.
%
Simulations are suitable tools to analyze systems that evolve with
time and involve randomness.
%
In general, simulations require less assumptions and simplifications
than theoretical approaches, and can be applied to systems where a
theoretical model can not be found.
%
A system described with the DE paradigm is characterized by a set of
actors and events, where actors interact causally through a series of
events on a time line and process these events in chronological order
\citep{ptolemaeus_system_2014, chung_simulation_2003,
ross_simulation_2012}.
%
In the case of study, the actors are the communicating civilizations
(CETIs), and the events are all meaningfull that produce a change in
the simulation variables, i.e., in the arrangement of CETIs and their
causal contacts.
%
Since it allows to explore several models easily, 
this method is well suited for the particular case of the difusion of
intelligence signals in the Galaxy.
%
Assuming some simple hypotheses, the discrete events method can be
performed taking into account a small number of variables, which
allow to analyze the variation of the results in the model parameter space.

@@

We also need to propose theoretical distribution functions for both the
distribution of lifespans and the distribution of the number of CETIs
per unit time \citep{maccone_evolution_2014,Sotos_biotechnology_2019}.

@@

The shapes of these distributions are discussed in Sec.~\ref{SS_PDF_shape}.

@@

As a consecuence of their finite lifetime,
each CETI will fill a spherical shell region of the Galaxy, limited by two
surfaces.

@@

as the squared distance from the source, imposes a
detectability limit.

@@

is covered with the emited signals.

@@

Any other CETI within this region will be aware of the originating CETI,
even if it has disapeared before the time of contact.

@@

This region will grow if the SFC has not reached $D_{max}$, and
will shrink otherwise, until the SLC reaches $D_{max}$.

@@

A that point, all signal from the central CETI is lost.

@@

In the 2D simulation only the intersection of these spheres with the
plane of the galaxy is relevant, and produce the corresponding
circles for the filled spheres or annular regions for the spherical
shells.

@@

We illustrate in the Fig.~\ref{F_scheme} the schematic
representation of the region causally connected to the central CETI.

@@

@@

@@

% Schematic representation of the growing communicating
%   sphere, over the space--time diagrams, where time is represented on
%   the horizontal direction, and space is represented in the vertical
%   direction.  In (a) the sphere is
%   growing as the surface of first contact has not reached the maximum
%   distance.  In (b) it has reached the maximum distance, so that it
%   remains at the same size.  After a Doomsday event, the signals can
%   still be observed, but the surface of last contact grows.
%   %
%   Schematic representation of two emitters, $E_1$ and $E_2$,
%   that reach each other at different times.  The time span for $E_i$
%   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
%   emitter $j$ between $C_{ij}$ and $B_{ij}$.
%   The type and length of causal contact in both directions depend on
%   the distance and time lag between the awakening events, the maximum
%   distance that a signal can reach and the time period in which each
%   emitter is active.

@@

@@

@@

The initially growing communicating sphere, is shown over
space--time diagrams, where time is represented on the horizontal
direction, and space is represented in the vertical direction.

@@

In the upper panel (a), the sphere is growing as the surface of first
contact has not reached the maximum distance.

@@

In the middle panel (b) it has reached the maximum distance, so that
it remains at the same size.

@@

After a Doomsday event (bottom panel, b), the signals can still be
observed, but the surface of last contact grows.

@@

We also show in this Figure a schematic representation of two
emitters, E$_1$ and E$_2$, that reach each other at different times.

@@

The time span for E$_i$ is (A$_i$ , D$_i$) , for $i = {1, 2}$.
Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
The type and length of causal contact in both directions depend on the
distance and time lag between the awakening events, the maximum
distance that a signal can reach and the time period in which each
emitter is active.
In the Fig.~\ref{F_messages} we show a schematic representation of the
possible cases in which a CETI A can be in causal contact with a CETI
B which appears later.
The duration of the causal contact in one direction depends on several
factors, mainly $\Delta$t~A , $\Delta$t~C , and $\Delta$X.

>>>>

Simulations are suitable tools to analyze systems that evolve with
time and involve randomness.
%
In general, simulations require less assumptions and simplifications
than theoretical approaches, and can be applied to systems where a
theoretical model can not be found.
%
Certain complex stochastic processes can be efficiently modeled with
the discrete--event (hereafter, DE) simulation approach.
%
A system described with the DE paradigm is characterized by a set of
actors and events, where actors interact causally through a series of
events on a time line and process these events in chronological order
\citep{ptolemaeus_system_2014, chung_simulation_2003,
ross_simulation_2012}.
%
We follow this approach to simulate the statistical properties of a
set of intelligent civilizations (CETIs) in the Galaxy that have the
ability of emitting and communicating messages at light speed.
%
Since this method to explore several models easily, it is well suited
for the particular case of the difusion of intelligence signals in the
Galaxy and its dependence on the statistical properties proposed for
the rise and fall of the communicating capacity in the galactic
intelligent civilizations.
%
The system is defined by a number of actors that represent CETIs and
appear at different instants in time, generating events that produce
meaningful changes in the variables that describe the system, i.e., in
the arrangement of CETIs and their network of causal contacts.
%
Assuming some simple hypotheses, the discrete events method can be
performed taking into account a small number of variables, which allow
to analyze the variation of the results in the model parameter space.

@@

We also need to propose theoretical distribution functions for both
the distribution of lifespans and the distribution of the number of
CETIs per unit time \citep{maccone_evolution_2014,
Sotos_biotechnology_2019}.

@@

The shapes of these distributions are discussed in
Sec.~\ref{SS_PDF_shape}.

@@

as the squared distance from the source, imposes a detectability
limit.
As a consecuence of this limitation and of their finite lifetime,
considered as the period of time between the aquisition and loss of
communicating capacity, each CETI will fill a spherical shell region
of the Galaxy, limited by two surfaces.

@@

@@

is filled with the emited signals.

@@

Any other CETI within this region will be aware of the originating
CETI, even if it has disapeared before the time of contact.

@@

This region will grow if the surface of first contact has not yet
reached the maximum distance $D_{max}$, and will shrink otherwise
until the surface of last contact reaches $D_{max}$, and producing as
a result the loss of all signal from the central CETI.

@@

In our approach, we consider a model Galaxy where the width of the
disk is negligigle with respect to the radius of the disk, so that
CETIs are placed on a 2D plane disk.

@@

In the 2D simulation only the intersection of the communicating
spherical shells with the plane of the galaxy is relevant, and produce
the corresponding circles or rings for the filled spheres or annular
regions of the spherical shells, respectively.

@@

We illustrate in the Fig.~\ref{F_scheme} the schematic representation
of the region causally connected to the central CETI.

@@

To this end, we use space--time diagrams, where time is represented on
the horizontal axis, and space is represented in the vertical axis.

@@

A growing surface of first contact in this diagram is represented as a
growing circle, centered on a point in the time axis, which shifts
towards the right at the same rate the radius of the circle increases.

@@

The maximum distance of the message limits the radius of the circles,
so that once this distance is reached, the circle continues its shift
towards the right, but maintaining a constant radius.

@@

At the end of the lifetime of a CETI, a growing surface of last
contact is represented by a circle that covers the inner part of the
causal region, producing a ring of diminishig area (panel (a)).

@@

In this scheme, the entire lifetime of a CETI can then be represented
by a polygon, limited in the spatial axis by the maximum distance of
the signal $D_{max}$ and in the time axis by the time span between the
starting time (concave vertex) and the ending time (convex vertex).

@@

This representation allows to easily visualize the important events
that result from the existence and communications of two CETIs (panel (b)).

@@

The points in time where the CETIs aquire their communicating
capacity (CC) are dubbed ''awakening'' events ($A_1$ and $A_2$ in the
Figure).

@@

Simmilarly, the points in time where the CETIs loss their communicating
capacity are dubbed ''doomsday'' events ($D_1$ and $D_2$ in the
Figure).

@@

The points in space--time where the first contact is produced for each
one of the CETIs, are defined as ''contact'' events, shown as $C_1$
and $C_2$ in the Figure.

@@

Finally, the two points where the contact is lost for each one the
CETIs, are denominated ''blackout'' events, shown as $B_1$ and $B_2$
in the Figure.

@@

It is important to notice that there is a time delay between the contact
events of the two CETIs involved in this analysis, and also a time
delay between the two blackout events, so that the time range when a
bidirectional contact is possible occurs between the maximum time of
the contact points and the minimum time between the blackout points.

%
% Schematic representation of the growing communicating
%   sphere, over the space--time diagrams, where time is represented on
%   the horizontal direction, and space is represented in the vertical
%   direction.  In (a) the sphere is
%   growing as the surface of first contact has not reached the maximum
%   distance.  In (b) it has reached the maximum distance, so that it
%   remains at the same size.  After a Doomsday event, the signals can
%   still be observed, but the surface of last contact grows.
%   %
%   Schematic representation of two emitters, $E_1$ and $E_2$,
%   that reach each other at different times.  The time span for $E_i$
%   is $(A_i, D_i)$, for $i=\{1,2\}$.  Emitter $i$ can listen to
%   emitter $j$ between $C_{ij}$ and $B_{ij}$.
%   The type and length of causal contact in both directions depend on
%   the distance and time lag between the awakening events, the maximum
%   distance that a signal can reach and the time period in which each
%   emitter is active.
%
%
%
The initially growing communicating sphere, is shown over
space--time diagrams, where time is represented on the horizontal
direction, and space is represented in the vertical direction.
%
In the upper panel (a), the sphere is growing as the surface of first
contact has not reached the maximum distance.
%
In the middle panel (b) it has reached the maximum distance, so that
it remains at the same size.
%
After a Doomsday event (bottom panel, b), the signals can still be
observed, but the surface of last contact grows.
%
We also show in this Figure a schematic representation of two
emitters, E$_1$ and E$_2$, that reach each other at different times.
%
The time span for E$_i$ is (A$_i$ , D$_i$) , for $i = {1, 2}$.
%
Emitter $i$ can listen to emitter $j$ between C$_{ij}$ and B$_{ij}$.
%
The type and length of causal contact in both directions depend on the
distance and time lag between the awakening events, the maximum
distance that a signal can reach and the time period in which each
emitter is active.
%
In the Fig.~\ref{F_messages} we show a schematic representation of the
possible cases in which a CETI A can be in causal contact with a CETI
B which appears later.
%
The duration of the causal contact in one direction depends on several
factors, mainly $\Delta$t~A , $\Delta$t~C , and $\Delta$X.
